---
title: "1.1 R REFRESHER"
output: html_document
---
# 1.1 R REFRESHER 

## QUESTION 1.1.1: R REFRESHER  (本题共有1分)
If you haven't done so already, install the library UsingR
```{r}
install.packages("UsingR")
```
Then once you load it you have access to Galton's father and son heights:
```{r}
library(UsingR)

?father.son
```
What is the average height of the sons (don't round off)?

```{r}
data(father.son)
summary(father.son$sheight)
mean(father.son$sheight)
```

## QUESTION 1.1.2 R REFRESHER  (本题共有1分)
One of the defining features of regression is that we stratify one variable based on others. In Statistics we use the verb "condition". For example, the linear model for son and father heights answers the question how tall do I expect a son to be if I condition on his father being x inches. The regression line answers this question for any x.

Using the father.son dataset described above, we want to know the expected height of sons if we condition on the father being 71 inches. Create a list of son height's for sons that have fathers with height of 71 inches (round to the nearest inch).

What is the mean of the son heights for fathers that have a height of 71 inches (don't round off your answer)? (Hint: use the function round() on the fathers' heights)

```{r}
names(father.son)
mean(father.son$sheight[round(father.son$fheight) == 71])
```

## QUESTION 1.2.1  (本题共有1分)
In R we have vectors and matrices. You can create your own vectors with the function c.
```{r}
c(1,5,3,4)
```
They are also the output of many functions such as
```{r}
rnorm(10)
```
You can turn vectors into matrices using functions such as rbind, cbind or matrix.

Create the matrix from the vector 1:1000 like this:
```{r}
X = matrix(1:1000,100,10)
```
What is the entry in row 25, column 3 ?

```{r}
X[25,3]
```

## QUESTION 1.2.2  (本题共有1分)
Using the function cbind, create a 10 x 5 matrix with first column
```{r}
x=1:10
```
Then columns 2*x, 3*x, 4*x and 5*x in columns 2 through 5.

What is the sum of the elements of the 7th row?

```{r}
M <- cbind(x, 2*x, 3*x, 4*x, 5*x)
sum(M[7,])
```

## QUESTION 1.2.3  (本题共有1分)
Which of the following creates a matrix with multiples of 3 in the third column?
```{r}
matrix(1:60,20,3)[,3]
matrix(1:60,20,3,byrow=TRUE)[,3]
x=11:20;rbind(x,2*x,3*x)[,3]
x=1:40;matrix(3*x,20,2)[,3]
```


## QUESTION 1.3.1  (本题共有1分)
Suppose X is a matrix in R. Which of the following is not equivalent to X?
```{r}
t( t(X) )
X %*% matrix(1,ncol(X) ) 
X*1 
X%*%diag(ncol(X))
```

## QUESTION 1.3.2  (本题共有1分)
Solve the following system of equations using R:
```{r}
3a + 4b - 5c + d = 10

2a + 2b + 2c - d = 5

a -b + 5c - 5d = 7

5a + d = 4
```
What is the solution for c:

```{r}
X = rbind(c(3,4,-5,1),
          c(2,2,2,-1),
          c(1,-1,5,-5),
          c(5,0,0,1))
Y = c(10,5,7,4)
solve(X)%*%Y
```

## QUESTION 1.4.1  (本题共有1分)

Load the following two matrices into R:
```{r}
a <- matrix(1:12, nrow=4)
b <- matrix(1:15, nrow=3)
```

Note the dimension of 'a' and the dimension of 'b'.

In the question below, we will use the matrix multiplication operator in R, %*%, to multiply these two matrices.

What is the value in the 3rd row and the 2nd column of the matrix product of 'a' and 'b'

```{r}
(a%*%b)[3,2]
```

## QUESTION 1.4.2  (本题共有1分)
Multiply the 3rd row of 'a' with the 2nd column of 'b', using the element-wise vector multiplication with *.

What is the sum of the elements in the resulting vector?

```{r}
sum(a[3,]*b[,2])
```

## QUESTION 1.5.1  (本题共有1分)
Suppose we are analyzing a set of 4 samples. The first two samples are from a treatment group A and the second two samples are from a treatment group B. This design can be represented with a model matrix like so:
```{r}
X <- matrix(c(1,1,1,1,0,0,1,1),nrow=4)
rownames(X) <- c("a","a","b","b")
```
X will then look like:

a 1 0
a 1 0
b 1 1
b 1 1

Suppose that the fitted parameters for a linear model give us:
```{r}
beta <- c(5, 2)
```
Use the matrix multiplication operator, %*%, in R to answer the following questions:


What is the fitted value for the A samples?
```{r}
X%*%beta
```


## QUESTION 1.5.2  (本题共有1分)
What is the fitted value for the B samples?
```{r}
X%*%beta
```

## QUESTION 1.5.3  (本题共有1分)
Suppose now we are comparing two treatments B and C to a control group A, each with two samples. This design can be represented with a model matrix like so:
```{r}
X <- matrix(c(1,1,1,1,1,1,0,0,1,1,0,0,0,0,0,0,1,1),nrow=6)
rownames(X) <- c("a","a","b","b","c","c")
```
which results in a matrix that looks like

a 1 0 0
a 1 0 0
b 1 1 0
b 1 1 0
c 1 0 1
c 1 0 1

Suppose that the fitted values for the linear model are given by:
```{r}
beta <- c(10,3,-3)
```
What is the fitted value for the B samples?
```{r}
X%*%beta
```

## QUESTION 1.5.4  (本题共有1分)
What is the fitted value for the C samples?
