---
title: "W2 Testing Linear Model"
output: html_document
---

The standard error of an estimate is the standard deviation of the sampling distribution of an estimate. In PH525.1x, we saw that our estimate of the mean of a population changed depending on the sample that we took from the population. If we repeatedly sampled from the population, and each time estimated the mean, the collection of mean estimates formed the sampling distribution of the estimate. When we took the standard deviation of those estimates, that was the standard error of our mean estimate.

Here, we aren't sampling individuals from a population, but we do have random noise in our observations Y. The estimate for the linear model terms (beta-hat) will not be the same if we were to re-run the experiment, because the random noise would be different. If we were to re-run the experiment many times, and estimate linear model terms (beta-hat) each time, this is called the sampling distribution of the estimates. If we take the standard deviation of all of these estimates from repetitions of the experiment, this is called the standard error of the estimate. While we are not sampling individuals, you can think about the repetition of the experiment that we are "sampling" new errors in our observation of Y.

# QUESTION 2.1.1  (本题共有1分)
We have shown how to find the least squares estimates with matrix algebra. These estimates are random variables as they are linear combinations of the data. For these estimates to be useful we also need to compute the standard errors.

Here we review standard errors in the context of linear models.

It is useful to think about where randomness comes from. In our falling object example, randomness was introduced through measurement errors. Every time we rerun the experiment a new set of measurement errors will be made which implies our data will be random. This implies that our estimate of, for example, the gravitational constant will change. The constant is fixed, but our estimates are not. To see this we can run a Monte Carlo simulation. Specifically we will generate the data repeatedly and compute the estimate for the quadratic term each time.
```{r}
g = 9.8 ## meters per second
h0 = 56.67
v0 = 0
n = 25
tt = seq(0,3.4,len=n) ##time in secs, t is a base function
y = h0 + v0 *tt - 0.5* g*tt^2 + rnorm(n,sd=1)
```

Now we act as if we didn't know h0, v0 and -0.5*g and use regression to estimate these. We can rewrite the model as y = b0 + b1 t + b2 t^2 + e and obtain the LSE we have used in this class. Note that g = -2 b2.

To obtain the LSE in R we could write:

```{r}
X = cbind(1,tt,tt^2)
A = solve(crossprod(X))%*%t(X)
```


Given how we have defined A, which of the following is the LSE of g, the acceleration due to gravity (suggestion: try the code in R)?