---
output: 
  html_document: 
    keep_md: yes
    self_contained: no
---
Lab 3 for Intorduction to computational finance
========================================================
# task 1 Covariance matrix
+ Compute the covariance between X and Y and assign it to sig_xy.
+ Construct the covariance matrix and assign it to Sigma_xy.
```{r}
# Standard deviations and correlation
sig_x <- 0.10
sig_y <- 0.05
rho_xy <- 0.9
# Covariance between X and Y
sig_xy <- rho_xy*sig_x*sig_y
# Covariance matrix
Sigma_xy <- matrix(c(sig_x^2, sig_xy, sig_xy, sig_y^2), nrow = 2)
```

# task 2 Simulate data
+ Simulate 100 observations from the bivariate normal distributed random variables X and Y. Use the covariance matrix Sigma_xy from the previous exercise and suppose that the means are ??X=0.05 and ??Y=0.025. Assign the result to xy_vals.
+ Have a look at the first few observations with the head() function.
```{r}
#install.packages("mvtnorm")
library(mvtnorm)
# The covariance matrix (Sigma_xy) is still in your workspace
# Means
mu_x <- 0.05
mu_y <- 0.025
# Simulate 100 observations
set.seed(123)  # for reproducibility
xy_vals <- rmvnorm(100, mean = c(mu_x, mu_y), sigma = Sigma_xy)
# Have a look at the first observations
head(xy_vals)
```

# task 3 Plot the simulated data
+ Create a scatterplot of the simulated observations
+ Customize the plot in the following way: specify the plot symbol with pch = 16, set the symbol size to 2, set the color to "blue", use the title "Bivariate normal: rho=0.9", and use the axis labels "x" and "y".
```{r}
# Create scatterplot
plot(xy_vals[,1], xy_vals[,2], pch = 16, cex = 2, col = "blue", main = "Bivariate normal: rho=0.9", ylab = "y", xlab = "x")
```

# task 4 Add lines to the plot
+ Add a vertical line that indicates the mean of X and a horizontal line that indicates the mean of Y.
```{r}
plot(xy_vals[,1], xy_vals[,2], pch = 16, cex = 2, col = "blue", main = "Bivariate normal: rho=0.9", ylab = "y", xlab = "x")
abline(h = mu_y)
abline(v = mu_x)
```

# task 5 Compute a joint probability
+ Compute the joint probability Pr(X???0,Y???0).
```{r}
plot(xy_vals[,1], xy_vals[,2], pch = 16, cex = 2, col = "blue", main = "Bivariate normal: rho=0.9", ylab = "y", xlab = "x")
# Add lines
abline(h = mu_y, v = mu_x)
# Add line segments
segments(x0 = 0, y0 = -1e10, x1 = 0, y1 = 0, col="red")
segments(x0 = -1e10, y0 = 0, x1 = 0, y1 = 0, col="red")
# Compute joint probability
pmvnorm(lower = -Inf, upper = 0, mean = c(mu_x, mu_y), sigma = Sigma_xy)
```

# task 6 Negatively correlated random variables
+ Change the code to perform the same analysis with negative correlation ??XY=???0.9 instead of ??XY=0.9.
```{r}
# Standard deviations and correlation
sig_x <- 0.10
sig_y <- 0.05
rho_xy <- -0.9
# Covariance between X and Y
sig_xy <- rho_xy * sig_x * sig_y
# Covariance matrix
Sigma_xy <- matrix(c(sig_x^2, sig_xy, sig_xy, sig_y^2), nrow = 2, ncol = 2)

# Means
mu_x <- 0.05
mu_y <- 0.025
# Simulate 100 observations
set.seed(123)  # for reproducibility
xy_vals <- rmvnorm(100, mean=c(mu_x, mu_y), sigma = Sigma_xy)
head(xy_vals)

# Create scatterplot
plot(xy_vals[, 1], xy_vals[, 2], pch = 16, cex = 2, col = "blue", 
     main = "Bivariate normal: rho=-0.9", xlab = "x", ylab = "y")
# Add lines
abline(h = mu_y, v = mu_x)
# Add line segments
segments(x0 = 0, y0 = -1e10, x1 = 0, y1 = 0, col="red")
segments(x0 = -1e10, y0 = 0, x1 = 0, y1 = 0, col="red")

# Compute joint probability
pmvnorm(lower = c(-Inf, -Inf), upper = c(0, 0), 
        mean = c(mu_x, mu_y), sigma = Sigma_xy)
```

# task 7 Uncorrelated random variables
+ Change the code to perform the same analysis with correlation ??XY=0 instead of ??XY=???0.9.
```{r}
# Standard deviations and correlation
sig_x <- 0.10
sig_y <- 0.05
rho_xy <- 0
# Covariance between X and Y
sig_xy <- rho_xy * sig_x * sig_y
# Covariance matrix
Sigma_xy <- matrix(c(sig_x^2, sig_xy, sig_xy, sig_y^2), nrow = 2, ncol = 2)

# Means
mu_x <- 0.05
mu_y <- 0.025
# Simulate 100 observations
set.seed(123)  # for reproducibility
xy_vals <- rmvnorm(100, mean=c(mu_x, mu_y), sigma = Sigma_xy)
head(xy_vals)

# Create scatterplot
plot(xy_vals[, 1], xy_vals[, 2], pch = 16, cex = 2, col = "blue", 
     main = "Bivariate normal: rho=0", xlab = "x", ylab = "y")
# Add lines
abline(h = mu_y, v = mu_x)
# Add line segments
segments(x0 = 0, y0 = -1e10, x1 = 0, y1 = 0, col="red")
segments(x0 = -1e10, y0 = 0, x1 = 0, y1 = 0, col="red")

# Compute joint probability
pmvnorm(lower = c(-Inf, -Inf), upper = c(0, 0), 
        mean = c(mu_x, mu_y), sigma = Sigma_xy)
```


# for quiz 3 computation part
```{r}
m <- matrix(c(0.1,0.1,0,0.2,0,0.1,0,0.2,0.3), nrow = 3, dimnames = list(c(1,2,3),c(1,2,3)))
X <- c(1,2,3)
Y <- c(1,2,3)
p_Y = colSums(m)
p_X = rowSums(m)
mu_Y = p_Y%*%Y
mu_X = p_X%*%X
var_X = (X-mu_X)^2 %*% p_X
var_Y = (Y-mu_Y)^2 %*% p_Y
sd_X = sqrt(var_X)
sd_Y = sqrt(var_Y)
Cov_XY <- 0 
for (row in 1:nrow(m)){
  for (col in 1:ncol(m)){
    Cov_XY <- Cov_XY + (X[row]-mu_X)*(Y[col]-mu_Y)*m[row,col]
  }
}
Cor_XY <- Cov_XY/sd_X/sd_Y

(exp(12*0.04+sqrt(12)*0.09*qnorm(0.05))-1)*100000
(exp(12*0.04+sqrt(12)*0.09*qnorm(0.01))-1)*100000
```

