Lab 4 for Intorduction to computational finance
========================================================
# task 1 Simulate data from a MA(1) model
+ Simulate 250 observations from the above MA(1) model with θ=0.5 and the constant 0.05. Assign the result to ma1_sim.
```{r}
set.seed(123);
# Simulate 250 observations from the described MA(1) model
ma1_sim <- arima.sim(model=list(ma=0.5), n=250, mean=0, sd=0.1) + 0.05
```

# task 2 Plot the data from the simulated MA(1) model
+ Make a line plot of the observations in ma1_sim with the title "MA(1) Process: mu=0.05, theta=0.5". Label the x-axis with "time" and the y-axis with "y(t)". A line plot can be specified by setting type="l" of the plot() function.
+ Add a horizontal line at zero.
```{r}
# A line plot of the simulated observations
plot(ma1_sim, type = "l", main = "MA(1) Process: mu=0.05, theta=0.5", xlab  = "time", ylab = "y(t)")
abline(h = 0)
```

# task 3 Plotting the theoretical and the sample ACF
+ Use the ARMAacf function to calculate the theoretical autocorrelations up to lag 10. Assign the result to acf_ma1_model.
+ Construct the plot of the sample autocorrelations up to lag 10 by using the acf() function. Set the title of the plot to "Sample ACF".
```{r}
# Generate the theoretical ACF with upto lag 10
acf_ma1_model <- ARMAacf(ma = 0.5, lag.max = 10)
# Split plotting window in three rows
par(mfrow=c(3,1))
# First plot: The simulated observations
plot(ma1_sim, type="l",main="MA(1) Process: mu=0.05, theta=0.5",xlab="time",ylab="y(t)")
abline(h=0)
# Second plot: Theoretical ACF
plot(1:10, acf_ma1_model[2:11], type="h", col="blue",  ylab="ACF", main="theoretical ACF")
# Third plot: Sample ACF
# Assign to tmp the Sample ACF
tmp <- acf(ma1_sim, lag.max = 10, main = "Sample ACF")
# Reset graphical window to only one graph
par(mfrow=c(1,1))
```

# task 4 A different MA(1) model
+ Now assume that θ=0.9 instead of θ=0.5. Rewrite the code on the right such that it simulates data from the correct model and displays the three graphs correctly.
```{r}
set.seed(123);
# Simulate 250 observations from the described MA(1) model
ma1_sim <- arima.sim(model=list(ma=0.9), n=250, mean=0, sd=0.1) + 0.05;
# Generate the theoretical ACF with upto lag 10
acf_ma1_model <- ARMAacf(ma=0.9, lag.max=10)
# Split plotting window in three rows
par(mfrow=c(3,1))
# First plot: The simulated observations
plot(ma1_sim, type="l",main="MA(1) Process: mu=0.05, theta=0.9",xlab="time",ylab="y(t)")
abline(h=0)
# Second plot: Theoretical ACF
plot(1:10, acf_ma1_model[2:11], type="h", col="blue", main="theoretical ACF")
# Third plot: Sample ACF
tmp = acf(ma1_sim, lag.max=10, main="Sample ACF")
# Reset graphical window to only one graph
par(mfrow=c(1,1))
```

# task 5 An AR(1) model
+ Assign to ar1_sim 250 observations simulated from the described AR(1) model.
+ Assign to acf_ar1_model the theoretical autocorrelations up to lag 10.
+ Create the three plot functions from the previous exercise but now for the AR(1) model instead of the MA(1) model. Have a look at the code of the previous exercise if you are unsure.
```{r}
set.seed(123);
# Simulate 250 observations from the described AR(1) model
ar1_sim <- arima.sim(model=list(ar=0.5), n=250, mean=0, sd=0.1) + 0.05;
# Generate the theoretical ACF with ten lags
acf_ar1_model <- ARMAacf(ar=0.5, lag.max=10)
# Split plotting window in three rows
par(mfrow=c(3,1))
# Generate the same three graphs as in the previous exercise 
par(mfrow=c(3,1))
# First plot: The simulated observations
plot(ar1_sim, type="l", main="AR(1) Process: mu=0.05, phi=0.5",xlab="time",ylab="y(t)")
abline(h=0)
# Second plot: Theoretical AFC
plot(1:10, acf_ar1_model[2:11], type="h", col="blue", main="theoretical ACF")
# Third plot: Sample AFC
tmp <- acf(ar1_sim, lag.max=10, main="Sample ACF")
# Reset plotting window to default
par(mfrow=c(1,1));
```

# Matrix Algebra Quiz
```{r}
A <- matrix(c(1,4,7,2,4,8,6,1,3), byrow = T, nrow = 3)
B <- matrix(c(4,4,0,5,9,1,2,2,5), byrow = T, nrow = 3)
x <- matrix(c(1,2,3), byrow = F, nrow = 3)
y <- matrix(c(5,2,7), byrow = F, nrow = 3)
```
# Question 1
```{r}
t(A)
```

# Question 2
```{r}
t(B)
```

# Question 3
```{r}
t(x)
```

# Question 4
```{r}
t(y)
```

# Question 5
```{r}
A+B
```

# Question 6
```{r}
A-B
```

# Question 7
```{r}
2*A
```

# Question 8
```{r}
A %*% x
```

# Question 9
```{r}
t(y) %*% A %*% x
```

# Question 10
```{r}
M <- matrix(c(1,1,2,4),byrow = T, nrow = 2)
b <- matrix(c(1,2),byrow = T, nrow = 2)
solve(M)%*%b
```

# Question 11
```{r}
x <- c(1/3,1/3,1/3)
mu <- c(0.01,0.04,0.02)
Sigma <- matrix(c(.1,.3,.1,.3,.15,-.2,.1,-.2,.08), nrow = 3)
t(mu)%*%x
```

# Question 12
```{r}
t(x)%*%Sigma%*%x
```


