Lab 6 for Intorduction to computational finance
========================================================
# Task 1 Download the data and calculate the returns
+ Make sure that you understand the code on the right and click 'Submit Answer' to run the code and start the lab.  
```{r ,warning=FALSE}
library(tseries)
library(PerformanceAnalytics)
VBLTX_prices <- get.hist.quote(instrument="vbltx", start="2005-09-01", end="2010-09-30", quote="AdjClose",provider="yahoo", origin="1970-01-01",compression="m", retclass="zoo", quiet = TRUE)
FMAGX_prices <- get.hist.quote(instrument="fmagx", start="2005-09-01", end="2010-09-30", quote="AdjClose",provider="yahoo", origin="1970-01-01",compression="m", retclass="zoo", quiet = TRUE)
SBUX_prices <- get.hist.quote(instrument="sbux", start="2005-09-01",end="2010-09-30", quote="AdjClose",provider="yahoo", origin="1970-01-01",compression="m", retclass="zoo", quiet = TRUE)

# Change the class of the time index to yearmon, which is appropriate for monthly data.
# index() and as.yearmon() are functions in the zoo package 
index(VBLTX_prices) <- as.yearmon(index(VBLTX_prices))
index(FMAGX_prices) <- as.yearmon(index(FMAGX_prices))
index(SBUX_prices) <- as.yearmon(index(SBUX_prices))

# Create merged price data
all_prices <- merge(VBLTX_prices, FMAGX_prices, SBUX_prices)
# Rename columns
colnames(all_prices) <- c("VBLTX", "FMAGX", "SBUX")

# Calculate cc returns as difference in log prices
all_returns <- diff(log(all_prices))

# Create matrix with returns
return_matrix <- coredata(all_returns)
```

# Task 2 The standard error of the variances
+ Assign to n_obs the number of observations in the matrix return_matrix.  
+ Assign to sigma2hat_vals the estimates of $σ_{i}^{2}$ for the three different assets.  
+ Assign to se_sigma2hat the estimates of $SE(σ^{2}_{i})$ and print the result to the console.  
```{r}
# The variable return_matrix is preloaded in your workspace

# Number of observations
n_obs <- nrow(return_matrix)

# Estimates of sigma2hat
sigma2hat_vals <- apply(return_matrix, MARGIN = 2, FUN = var)

# Standard Error of sigma2hat
se_sigma2hat <- sigma2hat_vals/sqrt(n_obs/2)

se_sigma2hat
```

# Task 3 Estimate the standard error of the correlation parameter

+ Assign to cor_matrix the correlation matrix of the returns.  
+ Assign to rhohat_vals the estimates of the correlations between "VBLTX,FMAGX","VBLTX,SBUX","FMAGX,SBUX".  
+ Assign to se_rhohat the estimates of $SE(\rho^{2}_{ij})$ and print the result to the console.
```{r}
# Calculate the correlation matrix
cor_matrix <- cor(return_matrix)

# Get the lower triangular part of that 'cor_matrix'
rhohat_vals <- cor_matrix[lower.tri(cor_matrix)]

# Set the names
names(rhohat_vals) <- c("VBLTX,FMAGX","VBLTX,SBUX","FMAGX,SBUX")

# Compute the estimated standard errors for correlation
se_rhohat <- (1-rhohat_vals^2)/sqrt(n_obs)

se_rhohat
```

# Task 4 Hypothesis test for the mean
+ Use the t.test to perform the t-test for μVBLTX and print the result to the console. What do you conclude?
```{r}
t.test(all_returns[,1],mu = 0)
```

# Task 5 Interpretation of the hypothesis test for the mean
```{r}
t.test(all_returns[,2],mu = 0)
```

# Task 6 Hypothesis test for the correlation
+ Use the cor.test() function to test whether the correlation between the returns of "VBLTX" and "FMAGX" is significantly different from zero. What do you conclude?
```{r}
cor.test(x = all_returns[,1], y = all_returns[,2])
```

# Task 7 Interpretation of the hypothesis test for correlation

```{r}
cor.test(x = all_returns[,1], y = all_returns[,3])

```

# Task 8 Normality of the asset returns
+ Use the jarque.bera.test to test the normality of the "VBLTX" returns and print the output to the console. What do you conclude?
```{r}
jarque.bera.test(all_returns[,1])
```

# Task 9 Interpretation of the normality test
```{r}
jarque.bera.test(all_returns[,2])
```

# Task 10 Bootstrapping
+ Use the boot() function to generate 999 bootstrap samples based on the returns of VBLTX and assign the result to VBLTX_mean_boot.
+ Plot the bootstrapped distribution of the mean and a QQ-plot. This can be done by applying the plot() function on the VBLTX_mean_boot that you have created in the previous step.
```{r}
library("boot")
# Function for bootstrapping sample mean: 
mean_boot <- function(x, idx) {
  ans <- mean(x[idx])
  ans 
} 
# Construct VBLTX_mean_boot:
VBLTX_mean_boot <- boot(return_matrix[,"VBLTX"], statistic = mean_boot, R=999)
# Print the class of VBLTX_mean_boot
class(VBLTX_mean_boot)

# Print VBLTX_mean_boot
VBLTX_mean_boot

# Plot bootstrap distribution and qq-plot against normal
plot(VBLTX_mean_boot)
```

