Lab 8 Introduction to Computational Finance
=======================================
# task 1 Loading in your data set
+ Load the working environment that contains the data set with the help of load().
+ Use head() and tail() to explore the returns_df data set.
+ The code chunks to create the timeplots of stocks on separate charts or within the same graph are provided. Make sure to have a close look at these.
```{r}

# Load the relevant packages
library("zoo")

# Load the working environment
load(url("http://s3.amazonaws.com/assets.datacamp.com/course/compfin/lab9.RData"))

# Explore the data set
head(returns_df)
tail(returns_df)

# Timeplots of stocks on seperate graphs
my.panel <- function(...) {
  lines(...)
  abline(h=0)
}
plot(returns_df, lwd=2, panel=my.panel, col="blue")

# Timeplots of stocks on same graph
plot(returns_df, plot.type = "single", main="Returns", col=1:4, lwd=2)
abline(h=0)
legend(x="bottomleft", legend=colnames(returns_df), col=1:4, lwd=2)
```


# Task 2 The CER model
+ Assign to sigma2_month the estimates of $\sigma^2_i$for all four assets.
+ Calculate sigma_month, the estimates of $$\sigma_i$ for all four assets.
+ Estimate the correlations $\rho_{ij}$ between all stocks, and assign the result to cor_mat_month.
+ Create the pairwise scatterplots between all four stocks. Use coredata() to extract the core data from returns_df and pairs() to create a matrix of scatter plots. Take the color blue and use 16 for points pch.
```{r}
# All data is preloaded in your workspace.  Type 'ls()' in the console to see what has been loaded.

# Parameters CER model
mu_hat_month <- apply(returns_df, 2, mean)
mu_hat_month
sigma2_month <- apply(returns_df, 2, var)
  
sigma2_month
sigma_month <- sigma2_month^0.5
  
sigma_month
cov_mat_month <- var(returns_df)
cov_mat_month
cor_mat_month <- cor(returns_df)
  
cor_mat_month

# Pairwise scatterplots
pairs(coredata(returns_df), pch = 16, col = "blue")  
```

# Task 3 The global minimum variance portfolio - Part One
+ Use the globalMin.portfolio() function to calculate the global minimum variance portfolio here.
+ Use plot() to visualize the weights of your four stocks in the global minimum variance portfolio.
```{r}
# All data and CER parameters are preloaded in your workspace.  Type 'ls()' in the console to see them.

# Calculate the global minimum variance portfolio
global_min_var_portfolio = globalMin.portfolio(er = mu_hat_month,
                    cov.mat = cov_mat_month,
                    shorts = T)

  
global_min_var_portfolio

# Plot the portfolio weights of our four stocks
plot(global_min_var_portfolio)
```

# Task 4 The global minimum variance portfolio - Part Two

+ The restriction matrices are already set up. Make sure to study their output.
+ Use these restriction matrices and the global minimum variance portfolio equality constraint to minimize the portfolio variance via solve.QP(). Assign the result to quad_prog.
```{r}
# All data and CER parameters are preloaded in your workspace.  Type 'ls()' in the console to see them.
library(quadprog)
# set restriction matrices
D_matrix <- 2* cov_mat_month
D_matrix
d_vector <- rep(0,4)
d_vector
A_matrix <- cbind(rep(1,4),diag(4))
A_matrix
b_vector <- c(1,rep(0,4))
b_vector

# use solve.QP to minimize portfolio variance
quad_prog <- solve.QP(Dmat=D_matrix,
                      dvec=d_vector,
                      Amat=A_matrix,
                      bvec=b_vector,
                      meq=1)
  
quad_prog
```

# Task 5 The global minimum variance portfolio - End game
+ Calculate the composition of the global minimum variance portfolio with no short sales allowed. This time using the globalMin.portfolio() instead of solve.QP() and the restriction matrices.
```{r}
# All data and CER parameters are preloaded in your workspace.  Type 'ls()' in the console to see them.

# The global minimum variance portfolio
global_min_var_portfolio <- globalMin.portfolio(er = mu_hat_month,
                    cov.mat = cov_mat_month,
                    shorts = F)
  
global_min_var_portfolio 
```

# Task 6 An efficient portfolio
+ Assign the highest average return to mu_target.
+ Use the efficient.portfolio() function to calculate your efficient portfolio characteristics for both cases, that is, one where short sales is allowed and one where short sales is not allowed. Assign the results to efficient_porfolio_short and efficient_porfolio_no_short.
+ Plot the weights of the stocks in both portfolios using the plot() function.
```{r}
# All data and CER parameters are preloaded in your workspace.  Type 'ls()' in the console to see them.

# highest average return
mu_target <- max(mu_hat_month)

# short sales allowed
efficient_porfolio_short <- efficient.portfolio(er = mu_hat_month,
                                                target.return = mu_target,
                                                cov.mat = cov_mat_month,
                                                shorts = T)
  
efficient_porfolio_short
plot(efficient_porfolio_short)

# no short sales allowed
efficient_porfolio_no_short <- efficient_porfolio_short <- efficient.portfolio(er = mu_hat_month,
                                                target.return = mu_target,
                                                cov.mat = cov_mat_month,
                                                shorts = F)
  
efficient_porfolio_no_short
plot(efficient_porfolio_no_short)
```

# Task 7 The efficient frontier
+ Use the efficient.frontier() function to calculate the expected returns and standard deviation of the portfolios lying on the curve. Make a grid for $\alpha$=1,0.9,…,−1. Assign the result to efficient_frontier. Type args(efficient.frontier) to print the arguments of that function to the console. Choose the appropriate values for alpha.min and alpha.max.
+ Make a plot of the efficient frontier curve via plot(). Set the colors of the dots to blue, the lwd argument to 2, and make sure that the stock names are displayed.
```{r}
# All data and CER parameters are preloaded in your workspace.  Type 'ls()' in the console to see them.

# The efficient frontier of risky assets 
efficient_frontier <- efficient.frontier(er = mu_hat_month,
                                         cov.mat = cov_mat_month,
                                         nport = 21,
                                         alpha.min = -1,
                                         alpha.max = 1,
                                         shorts = T)
  
summary(efficient_frontier)

# The plot
plot(efficient_frontier,plot.assets=T, col= "blue",lwd = 2)
```

# Task 8 The tangency portfolio
+ Assign the monthly risk-free rate to t_bill_rate.
+ Use the tangency.portfolio() function to calculate the efficient portfolio characteristics for both cases, that is, one where short sales is allowed, and one where short sales is not allowed. Assign the results to tangency_portfolio_short and tangency_portfolio_no_short.
+ Study the plots that represent the weights of the stocks in both portfolios. The corresponding plot() functions are already provided.
```{r}
# All data and CER parameters are preloaded in your workspace.  Type 'ls()' in the console to see them.

# risk free rate
t_bill_rate <- 0.005

# Tangency portfolio short sales allowed
tangency_portfolio_short <- tangency.portfolio(er = mu_hat_month,
                                               cov.mat =cov_mat_month,
                                               risk.free = t_bill_rate,
                                               shorts = T)
  
summary(tangency_portfolio_short)
#plot
plot(tangency_portfolio_short)

# Tangency portfolio short sales not allowed
tangency_portfolio_no_short <- tangency.portfolio(er = mu_hat_month,
                                               cov.mat =cov_mat_month,
                                               risk.free = t_bill_rate,
                                               shorts = F)
  
summary(tangency_portfolio_no_short)
#plot
plot(tangency_portfolio_no_short)
```

